---
title: "writing_functions"
author: "Juliet Rowe"
date: "2023-10-26"
output: github_document
---

```{r}
library(tidyverse)
library(rvest)
```


Set seed for reproducibility

```{r}
set.seed(12345)
```


### Z score function

Z scores subtract the mean ad divide by the sd

```{r}
x_vec = rnorm(20, mean=5, sd=0.3)
```

Compute Z scores for `x_vec`.

```{r}
(x_vec - mean(x_vec)) / sd(x_vec)
```


Write a function to do this!

```{r}
z_score = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument should be numbers")
  } else if (length(x) < 2) {
    stop("You need at least 2 numbers to get z scores")
  }
  
  z = (x_vec - mean(x)) / sd(x)
  
  z
  
}
```


Check that this works.

```{r}
z_score(x = x_vec)

z_score(x = rnorm(10, mean=5))
```


Keep checking.

```{r}


z_score(c("my", "name", "is", "jeff"))

z_score(c(TRUE, TRUE, FALSE, TRUE))

z_score(iris)
```


### Multiple outputs

Write a function that returns the mean and sd from a sample of numbers.

```{r}
mean_and_sd = function(x) {
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  
  
}
```


double check i did this right...

```{r}
z_score = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument should be numbers")
  } else if (length(x) < 2) {
    stop("You need at least 2 numbers to get z scores")
  }
  
   mean_x = mean(x)
   sd_x = sd(x)
  
  tibble(
    mean=mean_x,
    sd=sd_x
  )
  
}
```



### Start getting means and sds

```{r}
x_vec = rnorm(n = 30, mean = 5, sd = 0.5)

  tibble(
    mean=mean(x_vec),
    sd=sd(x_vec)
  )

```


Let's write a function that uses `n`, a true mean, and true SD as inputs. 

```{r}
sim_mean_sd = function(n_obs, mu, sigma) {
  
  x_vec = rnorm(n = n_obs, mean = mu, sd = sigma)

  tibble(
    mean=mean(x_vec),
    sd=sd(x_vec)
  )
}

sim_mean_sd(n_obs = 3000, mu=5, sigma=12.3)

sim_mean_sd(12, 24, 4)
```


## LoTR words.

```{r}
fellowship_ring = 
lotr_load_and_tidy = function(path, range, movie_name) {
  
  df = readxl::read_excel(path, range = range) |>
    janitor::clean_names() |>
    gather(key = sex, value = words, female:male) |>
    mutate(race = str_to_lower(race),
           movie = movie_name)
  
  df
  
}

lotr_tidy = 
  bind_rows(
    lotr_load_and_tidy("LotR_Words.xlsx", "B3:D6", "fellowship_ring"),
    lotr_load_and_tidy("LotR_Words.xlsx", "F3:H6", "two_towers"),
    lotr_load_and_tidy("LotR_Words.xlsx", "J3:L6", "return_king")) |>
  select(movie, everything()) 
```



### NSDUH

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```


Try to write a quick function

```{r}
nsduh_import = function(html, table_number, outcome_name){
  
  html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent), 
    outcome = outcome_name) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
}

nsduh_import(html = nsduh_html, table_number = 1, outcome_name = "marj")
nsduh_import(html = nsduh_html, table_number = 4, outcome_name = "cocaine")


```

